\textbf{COMP0086 Summative Assignment}

### 2022-11-15 submission.tex
add kernel ridge regression results

        \item The mean squared error for each fitted curve:
        \begin{center}
        \begin{tabular}{c|c|c|c|c}%
         \textbf{Metric}&\textbf{k=1}&\textbf{k=2}&\textbf{k=3} &\textbf{k=4}% specify table head
        \csvreader[head to column names]{outputs/q1/q1c.csv}{}% use head of csv as column names
        {\\\hline\csvcoli&\csvcolii&\csvcoliii&\csvcoliv&\csvcolv}% specify your coloumns here
        \end{tabular}
        \end{center}
    \end{enumerate}
\newpage
\item[2.]
    \begin{enumerate}
        \item
        \begin{enumerate}
            \item Plotting the function $sin^2(2\pi x)$ in the range $0 \leq x \leq 1$ with the sampled points superimposed:
                    \begin{figure}[h]
                    \centering
                    \includegraphics[scale=0.5]{outputs/q2/q2ai}
                    \caption{Noisy samples}
                    \label{fig:2ai}
                    \end{figure}
            \item Fitting the data set with polynomial bases of dimension $k = 2, 5, 10, 14, 18$ and plotting each of these 5 curves with the sampled points superimposed:
                    \begin{figure}[h]
                    \centering
                    \includegraphics[scale=0.5]{outputs/q2/q2aii}
                    \caption{Polynomial fit curves}
                    \label{fig:2aii}
                    \end{figure}
        \end{enumerate}
\newpage
        \item Plotting  the natural $\log$ ($\ln$) of the training error versus the polynomial dimension $k = 1, . . . , 18$:
            \begin{figure}[h]
            \centering
            \includegraphics[scale=0.5]{outputs/q2/q2b}
            \caption{Training Error of Polynomial Curves}
            \label{fig:2b}
            \end{figure}

        \item Plotting the $\ln$ of the test error for a single trial of 1000 test points versus the polynomial dimension $k = 1, . . . , 18$:
            \begin{figure}[h]
            \centering
            \includegraphics[scale=0.5]{outputs/q2/q2c}
            \caption{Testing Error of Polynomial Curves (1 trial)}
            \label{fig:2c}
            \end{figure}
\newpage
        \item Plotting the $\ln$ of the test error for 100 trials of 1000 test points versus the polynomial dimension $k = 1, . . . , 18$:
            \begin{figure}[h]
            \centering
            \includegraphics[scale=0.5]{outputs/q2/q2d}
            \caption{Testing Error of Polynomial Curves (100 trials)}
            \label{fig:2d}
            \end{figure}
    \end{enumerate}
\newpage
\item[3.] Repeating 2 (b-d) with $\sin$ basis functions:
            \begin{figure}[h]
            \centering
            \includegraphics[scale=0.5]{outputs/q3/q3b}
            \caption{Training Error of $\sin$ Curves}
            \label{fig:3b}
            \end{figure}
            \begin{figure}[h]
            \centering
            \includegraphics[scale=0.5]{outputs/q3/q3c}
            \caption{Testing Error of $\sin$ Curves (1 trial)}
            \label{fig:3c}
            \end{figure}
            \begin{figure}[h]
            \centering
            \includegraphics[scale=0.5]{outputs/q3/q3d}
            \caption{Testing Error of $\sin$ Curves (100 trials)}
            \label{fig:3d}
            \end{figure}
\newpage
\subsection{Filtered Boston housing and kernels}
    \item[4.]
    \begin{enumerate}
        \item 
        We interpret the naive linear regression model using the existing framework for least squares regressors.\\
        Namely, we have:\\
        $w^* = (X^TX)^{-1}X^{T} y, X = (1,\dots,1)^{T} $\\
        \[\implies w* = m^{-1}  \sum_{i} 1*y_{i} = \bar{y}\]\\
        \[\implies f(x_{i}) = 1 \times w^{*} = \bar{y}\]\\
        We split the data using our train test split function, and train naive, single and multiple linear regression models.
            We repeat this 20 times and report the MSE on training and test sets.\\

        Here are the plots of training and test error:\\
        \includegraphics[scale=0.5]{outputs/q4/q4_train.png}\\
        \includegraphics[scale=0.5]{outputs/q4/q4_test.png}\\
    \end{enumerate}
\newpage
\subsection{Kernelised ridge regression}
\item[5.]
We fit our kernel ridge regressors for each parameter pair over 5 cross validation folds,
and search for the pair minimising our mean squared error.\\

Here are the optimal parameters along with the minimal mean squared error:\\

\item The mean squared error for each fitted curve:
\begin{center}
\begin{tabular}{c|c|c}%
\textbf{Gamma}&\textbf{Sigma}&\textbf{MSE}% specify table head
\csvreader[head to column names]{outputs/q5/optimal_gkrr_parameters.csv}{}% use head of csv as column names
{\\\hline\csvcoli&\csvcolii&\csvcoliii}% specify your coloumns here
\end{tabular}
\end{center}

\\
Here is our contour plot for mse over our parameter grid:\\

\includegraphics[scale = 0.5]{outputs/q5/contour_plot.png}\\

\item[5.]

Here is our final regression report over all oour models, run 20 times each.\\

\item The mean squared error for each fitted curve:
\begin{center}
\begin{tabular}{c|c|c|c|c}%
\textbd{model}&\textbf{train mse}&\textbf{train std}&\textbf{test mse}&\textbf{test std}% specify table head
\csvreader[head to column names]{outputs/q5/regression_report.csv}{}% use head of csv as column names
{\\\hline\csvcoli&\csvcolii&\csvcoliii&\csvcoliv&\csvcolv}% specify your coloumns here
\end{tabular}
\end{center}

\section{PART II}
\subsection{$k$-Nearest Neighbors}
\item[6.]Generating a random $h$:
            \begin{figure}[h]
            \centering
            \includegraphics[scale=0.5]{outputs/q6/q6}
            \caption{A hypothesis $h_{S,v}$ visualised with $|S|=100$ and $v=3$}
            \label{fig:6a}
            \end{figure}
\newpage
\item[7.]
    \begin{enumerate}
        \item Producing a visualisation using Protocol A:
            \begin{figure}[h]
            \centering
            \includegraphics[scale=0.5]{outputs/q7/q7}
            \caption{Protocol A}
            \label{fig:7}
            \end{figure}
        \item We know that the data generating process uses $v=3$, so we would have initially expected the smallest test error to be located at the matching k-value, $k=3$.
        However, the data generating process for the training data set labels \textit{and} the testing data set labels were injected with noise, so k-values greater than 3 have a ``de-noising'' effect allowing for better performance at around $k=7$.
        As the k-value increases beyond $k=7$, this smoothing effect is too strong, causing the gradual increase in test error as the k value increases.
        For $k<7$ and in particular $k<v=3$ we see the error growing because the model has small $k$ and is able to fit to the noise injected into the data generating process, corrupting the model.
        By averaging over 100 runs, the plot is relatively smooth allowing us to make generalisations about the relationship between k-value and mean test set error.
    \end{enumerate}
\newpage
\item[8.]
    \begin{enumerate}
        \item Producing a visualisation using Protocol B:
            \begin{figure}[h]
            \centering
            \includegraphics[scale=0.5]{outputs/q8/q8}
            \caption{Protocol B}
            \label{fig:8}
            \end{figure}
        \item k-nearest neighbours is a non-parametric model, so when queried with a test point for prediction, the number of training points within a given fixed radius of that test point increases as the number of training points increases.
        If we consider this fixed radius `sufficiently close' to make an informed prediction, the number of points within this radius will increase with more training points.
        So as the number of training points increases, the optimal k would increase to include more of the training points within this same fixed radius.
        We can see in the plot above that we see this general trend as expected, however there is considerable noise in the trend and so a smoother plot would likely requires more runs.
    \end{enumerate}
\newpage
\section{PART III}
\subsection{Questions}
\item[9.] \begin{enumerate}
              \item let $K_{c}(\textbf{x},\textbf{z}) = c+ \sum_{i=1}^{n}x_{i}z_{i} = c + \textbf{x}^{T}\textbf{z}$\\
        Our function $K_{c}(\cdot, \cdot)$ is psd iff:\\
        $\forall \{ u_{i} \}_{1:m} \in \mathcal{R}, {\textbf{x}_{j}\}_{1:m} \in \mathcal{R}^{n}$:

>>>>

        \item The mean squared error for each fitted curve:
        \begin{center}
        \begin{tabular}{c|c|c|c|c}%
         \textbf{Metric}&\textbf{k=1}&\textbf{k=2}&\textbf{k=3} &\textbf{k=4}% specify table head
        \csvreader[head to column names]{outputs/q1/q1c.csv}{}% use head of csv as column names
        {\\\hline\csvcoli&\csvcolii&\csvcoliii&\csvcoliv&\csvcolv}% specify your coloumns here
        \end{tabular}
        \end{center}
    \end{enumerate}
\newpage
\item[2.]
    \begin{enumerate}
        \item
        \begin{enumerate}
            \item Plotting the function $sin^2(2\pi x)$ in the range $0 \leq x \leq 1$ with the sampled points superimposed:
                    \begin{figure}[h]
                    \centering
                    \includegraphics[scale=0.5]{outputs/q2/q2ai}
                    \caption{Noisy samples}
                    \label{fig:2ai}
                    \end{figure}
            \item Fitting the data set with polynomial bases of dimension $k = 2, 5, 10, 14, 18$ and plotting each of these 5 curves with the sampled points superimposed:
                    \begin{figure}[h]
                    \centering
                    \includegraphics[scale=0.5]{outputs/q2/q2aii}
                    \caption{Polynomial fit curves}
                    \label{fig:2aii}
                    \end{figure}
        \end{enumerate}
\newpage
        \item[(b \& c)] Plotting the $\ln$ of the train and test error for a single trial of 1000 test points versus the polynomial dimension $k = 1, . . . , 18$:
            \begin{figure}[h]
            \centering
            \includegraphics[scale=0.5]{outputs/q2/q2bc}
            \caption{Training and Testing Error of Polynomial Curves (1 trial)}
            \label{fig:2c}
            \end{figure}
        \item[(d)] Plotting the $\ln$ of the train and test error for 100 trials of 1000 test points versus the polynomial dimension $k = 1, . . . , 18$:
            \begin{figure}[h]
            \centering
            \includegraphics[scale=0.5]{outputs/q2/q2d}
            \caption{Training and Testing Error of Polynomial Curves (100 trials)}
            \label{fig:2d}
            \end{figure}
    \end{enumerate}
\newpage
\item[3.] Repeating 2 (b-d) with $\sin$ basis functions:
            \begin{figure}[h]
            \centering
            \includegraphics[scale=0.5]{outputs/q3/q3bc}
            \caption{Training and Testing Error of $\sin$ Curves}
            \label{fig:3b}
            \end{figure}
            \begin{figure}[h]
            \centering
            \includegraphics[scale=0.5]{outputs/q3/q3d}
            \caption{Testing and Testing Error of $\sin$ Curves (100 trials)}
            \label{fig:3d}
            \end{figure}
\newpage
\subsection{Filtered Boston housing and kernels}
    \item[4.]
    \begin{enumerate}
        \item[(a)]
        We interpret the naive linear regression model using the existing framework for least squares regressors.\\
        Namely, we have:\\
        $w^* = (X^TX)^{-1}X^{T} y, X = (1,\dots,1)^{T} $\\
        \[\implies w* = m^{-1}  \sum_{i} 1*y_{i} = \bar{y}\]\\
        \[\implies f(x_{i}) = 1 \times w^{*} = \bar{y}\]\\
        Our model in this case is essentially always predicting the average housing price in the training data set.
%        We split the data using our train test split function, and train naive, single and multiple linear regression models.
%            We repeat this 20 times and report the MSE on training and test sets.\\

        \item[(b-d)] Here is a table of the training and test mean squared errors:\\
        \begin{center}
        \scalebox{0.8} {\begin{tabular}{c|c|c|c|c}%
         \textbf{Model}&\textbf{Train Mean}&\textbf{Train St. Dev.}&\textbf{Test Mean} &\textbf{Test St. Dev.}% specify table head
        \csvreader[head to column names]{outputs/q4/performance.csv}{}% use head of csv as column names
        {\\\hline\csvcoli&\csvcolii&\csvcoliii&\csvcoliv&\csvcolv}% specify your coloumns here
        \end{tabular}
        }
        \end{center}
    \end{enumerate}
\newpage
\subsection{Kernelised ridge regression}
\item[5.]
%We fit our kernel ridge regressors for each parameter pair over 5 cross validation folds,
%and search for the pair minimising our mean squared error.\\

%Here are the optimal parameters along with the minimal mean squared error:\\

%The mean squared error for each fitted curve:
%\begin{center}
%\begin{tabular}{c|c|c}%
%\textbf{Gamma}&\textbf{Sigma}&\textbf{MSE}% specify table head
%\csvreader[head to column names]{outputs/q5/optimal_gkrr_parameters.csv}{}% use head of csv as column names
%{\\\hline\csvcoli&\csvcolii&\csvcoliii}% specify your coloumns here
%\end{tabular}
%\end{center}
%
%\\

    \begin{enumerate}
\item[(a)] The optimal parameters:

    \begin{center}
    \scalebox{0.8} {\begin{tabular}{c|c|c|c|c}%
     \textbf{Trial}&$\log_2(\sigma)$&$\log_2(\gamma)$ % specify table head
    \csvreader[head to column names]{outputs/q5/q5a-optimal-params.csv}{}% use head of csv as column names
    {\\\hline\csvcolii&\csvcoliii&\csvcoliv}% specify your coloumns here
    \end{tabular}
    }
    \end{center}

\item[(b)] Displaying the logarithm of the mean 5 fold mean squared error for different parameters for a single trial:\\

\begin{centering}
\includegraphics[scale = 0.5]{outputs/q5/q5b-cross-valid-error}\\
\end{centering}
\item[(c)]The mean squared error for the training and test sets for the best $\gamma$ and $\sigma$ for a single trial:
    \begin{center}
    \scalebox{0.8} {\begin{tabular}{c|c|c|c|c}%
     \textbf{Model}&\textbf{Train Mean}&\textbf{Test Mean} % specify table head
    \csvreader[head to column names]{outputs/q5/q5c-performance.csv}{}% use head of csv as column names
    {\\\hline\csvcoli&\csvcolii&\csvcoliv}% specify your coloumns here
    \end{tabular}
    }
    \end{center}

\item[(d)]The mean squared error for each fitted curve over 20 runs:
        \begin{center}
        \scalebox{0.8} {\begin{tabular}{c|c|c|c|c}%
         \textbf{Model}&\textbf{Train Mean}&\textbf{Train St. Dev.}&\textbf{Test Mean} &\textbf{Test St. Dev.}% specify table head
        \csvreader[head to column names]{outputs/q5/q5d-performance.csv}{}% use head of csv as column names
        {\\\hline\csvcoli&\csvcolii&\csvcoliii&\csvcoliv&\csvcolv}% specify your coloumns here
        \end{tabular}
        }
        \end{center}
    \end{enumerate}
\newpage
\section{PART II}
\subsection{$k$-Nearest Neighbors}
\item[6.]Generating a random $h$:
            \begin{figure}[h]
            \centering
            \includegraphics[scale=0.5]{outputs/q6/q6}
            \caption{A hypothesis $h_{S,v}$ visualised with $|S|=100$ and $v=3$}
            \label{fig:6a}
            \end{figure}
\newpage
\item[7.]
    \begin{enumerate}
        \item Producing a visualisation using Protocol A:
            \begin{figure}[h]
            \centering
            \includegraphics[scale=0.5]{outputs/q7/q7}
            \caption{Protocol A}
            \label{fig:7}
            \end{figure}
        \item We know that the data generating process uses $v=3$, so we would have initially expected the smallest test error to be located at the matching k-value, $k=3$.
        However, the data generating process for the training data set labels \textit{and} the testing data set labels were injected with noise, so k-values greater than 3 have a ``de-noising'' effect allowing for better performance at around $k=7$.
        As the k-value increases beyond $k=7$, this smoothing effect is too strong, causing the gradual increase in test error as the k value increases.
        For $k<7$ and in particular $k<v=3$ we see the error growing because the model has small $k$ and is able to fit to the noise injected into the data generating process, corrupting the model.
        By averaging over 100 runs, the plot is relatively smooth allowing us to make generalisations about the relationship between k-value and mean test set error.
    \end{enumerate}
\newpage
\item[8.]
    \begin{enumerate}
        \item Producing a visualisation using Protocol B:
            \begin{figure}[h]
            \centering
            \includegraphics[scale=0.5]{outputs/q8/q8}
            \caption{Protocol B}
            \label{fig:8}
            \end{figure}
        \item k-nearest neighbours is a non-parametric model, so when queried with a test point for prediction, the number of training points within a given fixed radius of that test point increases as the number of training points increases.
        If we consider this fixed radius `sufficiently close' to make an informed prediction, the number of points within this radius will increase with more training points.
        So as the number of training points increases, the optimal k would increase to include more of the training points within this same fixed radius.
        We can see in the plot above that we see this general trend as expected, however there is considerable noise in the trend and so a smoother plot would likely requires more runs.
    \end{enumerate}
\newpage
\section{PART III}
\subsection{Questions}
\item[9.] \begin{enumerate}
              \item let $K_{c}(\textbf{x},\textbf{z}) = c+ \sum_{i=1}^{n}x_{i}z_{i} = c + \textbf{x}^{T}\textbf{z}$\\
        Our function $K_{c}(\cdot, \cdot)$ is psd iff:\\
        $\forall \{ u_{i} \}_{1:m} \in \mathcal{R}, {\textbf{x}_{j}\}_{1:m} \in \mathcal{R}^{n}$:
